---
title: "Assignment4_Q3"
output: html_document
date: "2024-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load necessary libraries
library(qicharts2)
library(dplyr)
library(lubridate)
```

```{r}
# Read the CSV file
data <- read.csv("monitor.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r}
# Convert Timestamp to Date object and add Day of the Year
data$Timestamp <- as.POSIXct(data$Timestamp, format = "%Y-%m-%d %H:%M:%S")
data$DayOfYear <- yday(data$Timestamp)
data
```

The data is started from 1st February 2021, so the DayOfYear column starts withe the value of 32.

## a) Is the memory usage of the server in control?

```{r}
a <- qicharts2::qic(x = DayOfYear, y = ProcessMemory, data = data, chart = "xbar", subtitle = "Experiment 1",freeze = 40)
plot(a)
```

```{r}
summary(a)
```

```{r}
summary(a) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 1 and sigma.signal=15 of this xbar chart summary suggests that the mean has a problem.

```{r}
b <- qicharts2::qic(x = DayOfYear, y = ProcessMemory, data = data, chart = "s", subtitle = "Experiment 2",freeze = 40)
plot(b)
```

```{r}
summary(b)
```

```{r}
summary(b) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 0 and sigma.signal=0 of this S chart summary suggests that the standard deviation doesn't have a problem.

From the analysis of these plots, we can conclude that there is a noticeable drift in the mean, while the spread(standard deviation) remains stable. This indicates that the memory usage on the server is **not in control**.

## b) Is the prediction time of the model in control?

```{r}
aa <- qicharts2::qic(x = DayOfYear, y = PredictionTimeMS, data = data, chart = "xbar", subtitle = "Experiment 3",freeze = 40)
plot(aa)
```

```{r}
summary(aa)
```

```{r}
summary(aa) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 1 and sigma.signal=4 of this xbar chart summary suggests that the mean has a problem.

```{r}
bb <- qicharts2::qic(x = DayOfYear, y = PredictionTimeMS, data = data, chart = "s", subtitle = "Experiment 4",freeze = 40)
plot(bb)
```

```{r}
summary(bb)
```

```{r}
summary(bb) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 1 and sigma.signal=14 of this S chart summary suggests that the standard deviation has a problem.

From the analysis of these plots, we can conclude that there is a noticeable drift in the mean, and a noticeable drift in the spread(standard deviation) . This indicates that the prediction time of the model is **not in control**.

## c) Is the stream of predictions in control?

```{r}
aaa <- qicharts2::qic(x = DayOfYear, y = Prediction, data = data, chart = "xbar", subtitle = "Experiment 5",freeze = 40)
plot(aaa)
```

```{r}
summary(aaa)
```

```{r}
summary(aaa) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 1 and sigma.signal=0 of this xbar chart summary suggests that the mean has a problem. Here although the points are within the range, there is one instance where the longest run is greater than the longest run max.

```{r}
bbb <- qicharts2::qic(x = DayOfYear, y = Prediction, data = data, chart = "s", subtitle = "Experiment 6",freeze = 40)
plot(bbb)
```

```{r}
summary(bbb)
```

```{r}
summary(bbb) %>%
  select(runs.signal, sigma.signal)
```

The run.signal = 1 and sigma.signal=0 of this S chart summary suggests that the standard deviation has a problem. Here, although the points are within the range, there is one instance where the number of crossings are less than the minimum number of crossing value.

From the analysis of these plots, we can conclude that there can be a drift in the mean, and a in the spread(standard deviation) . This indicates that the prediction time of the model is **not in control**.
